---
title: "Desmid Mini Project - ANOVA"
author: "Kristine Gesulga"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    code_folding: "hide" 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, include = FALSE, warnings = FALSE}

rm(list=ls())

# Load libraries
library(tidyverse) # lots of useful functions!
library(calecopal) # pretty, California-inspired palettes 
library(cowplot) # to arrange plots with labels
library(gridGraphics) # to span title across multiple plots
library(AICcmodavg) # to test AICc
library(car) # for Levene's Test

# Load functions
source("functions/flowcam_plotfunctions.R")

# Load data
raw_dat <- read_csv("data/1b_Growth_community_KG_edit2_mumax.csv")
glimpse(raw_dat)
flowcam_dat <- read_csv("data/flowcam_dat.csv")
glimpse(flowcam_dat)

```

# Community relative abundance
*Question: How does relative abundance change over time? Under nutrient-limitation? In a warmer temperature?*

<details>
  <summary>Click for details on tidying data.</summary>
```{r}
dat4plots <- flowcam_dat %>% 
  select(day, community, media, temp, species, sp_mean, total_desmids_mean, rel_abund) %>% 
  mutate(day_txt = case_when(day == 0 ~ "Day 0",
                             day == 27 ~ "Day 27")) %>% 
  mutate(species_txt = case_when(species == "e_ansatum" ~ "Euastrum ansatum",
                                 species == "s_teliferum" ~ "Staurastrum teliferum",
                                 species == "n_digitus" ~ "Netrium digitus",
                                 species == "s_gracile" ~ "Staurastrum gracile",
                                 species == "e_elegans" ~ "Euastrum elegans",
                                 species == "e_oblongum" ~ "Euastrum oblongum")) %>%   
  mutate(media_txt = case_when(media == "WHM" ~ "WHM",
                               media == "5N" ~ "WHM [5% N]",
                               media == "5P" ~ "WHM [5% P]"))
glimpse(dat4plots)

```
</details>

## Meso community

```{r}
# rel abund
pt0 <- plot_relabund(timestep_day = 0, comm = "meso", palette = "lake") #+
  # theme(legend.position="left",
  #       legend.text = element_text(size=6)) #c(.5,0)) 
pt0

pt27 <- plot_relabund(timestep_day = 27, comm = "meso", palette = "lake") #+
  # theme(legend.position = "none")
pt27

```


```{r}
p1 <- dat4plots %>% 
    group_by(day_txt, community, media, temp, species, rel_abund) %>% 
    distinct(rel_abund, .keep_all = T) %>% 
    ungroup() %>% 
    filter(day == 27,
           species == "e_elegans") %>% # add in later filter == community
    # plot
    ggplot(aes(fill=species_txt, y=rel_abund, x=as.factor(temp))) + 
    geom_bar(colour="black", position="stack", stat="identity") +
    facet_wrap(~media_txt) + #, nrow = 2, ncol = 3
    scale_fill_manual(values = "#CECEB9") + #cal_palette("lake")
    ylab("Relative abundance [%]") +
    xlab("Temperature [°C]") +
    labs(fill = "Species",
         # change here the community with func argument
         title = paste("Day", 27),
         subtitle = "Euastrum elegans (meso)") + 
    scale_y_continuous(expand = c(0,0)) +
    theme_bw() +
    # Customize facet labels
    theme(strip.background = element_rect(fill="#E4DECE"),
          strip.text.x = element_text(size = 10.5),
          legend.position = "none")

p2<- dat4plots %>% 
    group_by(day_txt, community, media, temp, species, rel_abund) %>% 
    distinct(rel_abund, .keep_all = T) %>% 
    ungroup() %>% 
    filter(day == 27,
           species == "e_oblongum") %>% # add in later filter == community
    # plot
    ggplot(aes(fill=species_txt, y=rel_abund, x=as.factor(temp))) + 
    geom_bar(colour="black", position="stack", stat="identity") +
    facet_wrap(~media_txt) + #, nrow = 2, ncol = 3
    scale_fill_manual(values = "#7AC9B7") + #cal_palette("lake")
    ylab("Relative abundance [%]") +
    xlab("Temperature [°C]") +
    labs(fill = "Species",
         # change here the community with func argument
         title = paste("Day", 27),
         subtitle = "Euastrum oblongum (meso)") + 
    scale_y_continuous(expand = c(0,0)) +
    theme_bw() +
    # Customize facet labels
    theme(strip.background = element_rect(fill="#E4DECE"),
          strip.text.x = element_text(size = 10.5),
          legend.position = "none")

p3 <- dat4plots %>% 
    group_by(day_txt, community, media, temp, species, rel_abund) %>% 
    distinct(rel_abund, .keep_all = T) %>% 
    ungroup() %>% 
    filter(day == 27,
           species == "s_gracile") %>% # add in later filter == community
    # plot
    ggplot(aes(fill=species_txt, y=rel_abund, x=as.factor(temp))) + 
    geom_bar(colour="black", position="stack", stat="identity") +
    facet_wrap(~media_txt) + #, nrow = 2, ncol = 3
    scale_fill_manual(values = "#6CA184") + #cal_palette("lake")
    ylab("Relative abundance [%]") +
    xlab("Temperature [°C]") +
    labs(fill = "Species",
         # change here the community with func argument
         title = paste("Day", 27),
         subtitle = "Staurastrum gracile (meso)") + 
    scale_y_continuous(expand = c(0,0)) +
    theme_bw() +
    # Customize facet labels
    theme(strip.background = element_rect(fill="#E4DECE"),
          strip.text.x = element_text(size = 10.5),
          legend.position = "none")

btmrow <- plot_grid(p1, p2, p3, labels = c('A', 'B', 'C'), label_size = 12, ncol = 2)

# now add the title
title <- ggdraw() +
  draw_label(
    "Relative abundance of individual species in mesotrophic community",
    fontface = 'bold',
    x = 0,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )

plot_grid(
  title, btmrow,
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1)
)
```

<br>

## Meso-oligo community
add plots tomorrow

# ANOVA

## Maximum growth rate
*Question a: Does temperature and/or nutrients have a significant effect on maximum growth rate (mumax)?*<br>
*Question b: Do temperature and nutrients have an interactive effect?*

### Meso community
<details>
  <summary>Click for details on tidying data.</summary>
```{r, message = FALSE}
dat_meso <- raw_dat %>% 
  filter(community == "meso") %>% 
  select(mumax, Temp, Nutrient)
glimpse(dat_meso) # response variable = mumax; explanatory variable = temperature / nutrient

# Convert Temperature + Nutrient treatments to factors (ANOVA test likes it in this format)
dat_meso$Temp <- as.factor(dat_meso$Temp)
dat_meso$Nutrient <- as.factor(dat_meso$Nutrient)
# Sanity check
table(dat_meso$Temp, dat_meso$Nutrient)
```
</details> 

<br>

#### Visualise the data

```{r}
# Plot
ggplot(dat_meso, aes(x = Nutrient, y = mumax, color = Temp)) +
  geom_boxplot()+
  labs(title = "Meso community")
```

<br>

<details>
  <summary>Click for details on model fitting, AICc test, and anova assumptions. </summary>
<br>
**Fit model**
```{r, message = FALSE}
# Additive model
meso_model_add <- lm(mumax ~ Temp + Nutrient, data = dat_meso)
meso_model_add

# Interactive model
meso_model_interx <- lm(mumax ~ Temp*Nutrient, data = dat_meso)
meso_model_interx

```

<br>

**Test AICc**
```{r}
models <- list(meso_model_add, meso_model_interx) 
model.names <- c("meso_model_add", "meso_model_interx")

aictab(cand.set = models, modnames = model.names)

```

*Based on AICc test, the additive model is a better fit.*

<br>

**Check assumptions of interactive model**

```{r}
# 1. Residuals-Fit plot check that variance is constant across groups
# --> Want: straight line; Avoid: funnel-like shape 
# 2. QQ plot checks for normality in the residuals
# --> Want: points fitting closely to the diagonal line
# https://mgimond.github.io/Stats-in-R/ANOVA.html

# Normality
OP <- par(mfrow=c(1,2))
plot(meso_model_interx, 1:2)
par(OP)

# Homogeneity of Variance
leveneTest(mumax~Temp*Nutrient, data = dat_meso)
```

*Data meets assumptions for the most part.*
</details>

<br>


#### Run ANOVA {.tabset}
##### Interactive ANOVA
```{r}
anova(meso_model_interx)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 2px; padding: 5px;}
</style>
<div class = "blue">

There is no significant effect of temp and nutrients on mumax or interactive effects.

</div>

<br><br><br>

##### Additive ANOVA
```{r}
anova(meso_model_add)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 2px; padding: 5px;}
</style>
<div class = "blue">

There is no significant effect of temp and nutrients on mumax.

</div>

<br><br><br>

### Meso-oligo community
<details>
  <summary>Click for details on tidying data.</summary>
```{r}
dat_mo <- raw_dat %>% 
  filter(community == "meso-oligo") %>% 
  select(mumax, Temp, Nutrient)
glimpse(dat_mo) # response variable = mumax; explanatory variable = temperature / nutrient

#### Convert Temperature + Nutrient treatments to factors (ANOVA test likes it in this format)
dat_mo$Temp <- as.factor(dat_mo$Temp)
dat_mo$Nutrient <- as.factor(dat_mo$Nutrient)
#### Sanity check
table(dat_mo$Temp, dat_mo$Nutrient)
```
</details> 

<br>

#### Visualise the data

```{r}
#### Plot
ggplot(dat_mo, aes(x = Nutrient, y = mumax, color = Temp)) +
  geom_boxplot()+
  labs(title = "Meso-oligo community")
```

<br>

<details>
  <summary>Click for details on model fitting, AICc test, and anova assumptions. </summary>
<br>
**Fit model**
```{r}
# Additive model
mo_model_add <- lm(mumax ~ Temp + Nutrient, data = dat_mo)
mo_model_add

# Interactive model
mo_model_interx <- lm(mumax ~ Temp*Nutrient, data = dat_mo)
mo_model_interx

```

<br>
**Test AICc**
```{r}
models <- list(mo_model_add, mo_model_interx) 
model.names <- c("mo_model_add", "mo_model_interx")

aictab(cand.set = models, modnames = model.names)

```

*Based on AICc test, the additive model is a better fit.*

<br>

**Check assumptions of interactive model**
```{r}
# Normality
OP <- par(mfrow=c(1,2))
plot(mo_model_interx, 1:2)
par(OP)

# Homogeneity of Variance
leveneTest(mumax~Temp*Nutrient, data = dat_mo)
```

*Data meets assumptions for the most part.*

</details>

<br>

#### Run ANOVA {.tabset}
##### Interactive ANOVA
```{r}
anova(mo_model_interx)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 2px; padding: 5px;}
</style>
<div class = "blue">

Temp and nutrients do not have a significant effect on mumax; however, there is a significant interaction effect between them (p < 0.1). 

</div>


##### Additive ANOVA
```{r}
anova(mo_model_add)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 2px; padding: 5px;}
</style>
<div class = "blue">

There is no significant effect of temp and nutrients on mumax.

</div>


## Average growth rate 
in progress...


# Session Info

<details>
  <summary>Click for details</summary>
```{r}
sessionInfo()
```
</details>
<br><br><br>
#endscript