---
title: "Desmid Mini Project - ANOVA"
author: "Kristine Gesulga"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, include = FALSE, warnings = FALSE}

rm(list=ls())

# Load libraries
library(tidyverse) # lots of useful functions!
library(AICcmodavg) # to test AIC
library(car) # for Levene's Test

# Load data
setwd("C:/Users/User/Desktop/Desmids/4_analyses")
raw_dat <- read_csv("1b_Growth_community_KG_edit2_mumax.csv")
glimpse(raw_dat)

```

# Maximum growth rate
Question 1a: Does temperature and/or nutrients have a significant effect on maximum growth rate (mumax)?<br>
Question 1b: Do temperature and nutrients have an interactive effect?

## Meso community
### Tidy data

<details>
  <summary>Click for details</summary>
```{r}
dat_meso <- raw_dat %>% 
  filter(community == "meso") %>% 
  select(mumax, Temp, Nutrient)
glimpse(dat_meso) # response variable = mumax; explanatory variable = temperature / nutrient

#### Convert Temperature + Nutrient treatments to factors (ANOVA test likes it in this format)
dat_meso$Temp <- as.factor(dat_meso$Temp)
dat_meso$Nutrient <- as.factor(dat_meso$Nutrient)
#### Sanity check
table(dat_meso$Temp, dat_meso$Nutrient)
```
</details> 


### Visualise the data

```{r, echo = FALSE}
#### Plot
ggplot(dat_meso, aes(x = Nutrient, y = mumax, color = Temp)) +
  geom_boxplot()+
  labs(title = "meso community")
```

### Fit model

```{r}
# Additive model
meso_model_add <- lm(mumax ~ Temp + Nutrient, data = dat_meso)
meso_model_add

# Interactive model
meso_model_interx <- lm(mumax ~ Temp*Nutrient, data = dat_meso)
meso_model_interx

```

### Test AICc

```{r}
models <- list(meso_model_add, meso_model_interx) 
model.names <- c("meso_model_add", "meso_model_interx")

aictab(cand.set = models, modnames = model.names)

```

> Based on AICc test, the additive model is a better fit.


### Check assumptions {.tabset}
#### Interactive model

```{r, echo = FALSE}
# 1. Residuals-Fit plot check that variance is constant across groups
# --> Want: straight line; Avoid: funnel-like shape 
# 2. QQ plot checks for normality in the residuals
# --> Want: points fitting closely to the diagonal line
# https://mgimond.github.io/Stats-in-R/ANOVA.html

# Normality
OP <- par(mfrow=c(1,2))
plot(meso_model_interx, 1:2)
par(OP)

# Homogeneity of Variance
leveneTest(mumax~Temp*Nutrient, data = dat_meso)
```

> Data meets assumptions for the most part.


#### Additive model
```{r, echo = FALSE}
# plot(meso_model_interx, 1)
# plot(meso_model_interx, 2)
OP <- par(mfrow=c(1,2))
plot(meso_model_add, 1:2)
par(OP)
```




### Run ANOVA {.tabset}
#### Interactive ANOVA
```{r}
anova(meso_model_interx)
```
> There is no significant effect of temp and nutrients on mumax or interactive effects.

#### Additive ANOVA
```{r}
anova(meso_model_add)
```

> There is no significant effect of temp and nutrients on mumax.

</details>


## Meso-oligo community
### Tidy data

<details>
  <summary>Click for details</summary>
```{r}
dat_mo <- raw_dat %>% 
  filter(community == "meso-oligo") %>% 
  select(mumax, Temp, Nutrient)
glimpse(dat_mo) # response variable = mumax; explanatory variable = temperature / nutrient

#### Convert Temperature + Nutrient treatments to factors (ANOVA test likes it in this format)
dat_mo$Temp <- as.factor(dat_mo$Temp)
dat_mo$Nutrient <- as.factor(dat_mo$Nutrient)
#### Sanity check
table(dat_mo$Temp, dat_mo$Nutrient)
```
</details> 


### Visualise the data

```{r, echo = FALSE}
#### Plot
ggplot(dat_mo, aes(x = Nutrient, y = mumax, color = Temp)) +
  geom_boxplot()+
  labs(title = "meso-oligo community")
```

### Fit model

```{r}
# Additive model
mo_model_add <- lm(mumax ~ Temp + Nutrient, data = dat_mo)
mo_model_add

# Interactive model
mo_model_interx <- lm(mumax ~ Temp*Nutrient, data = dat_mo)
mo_model_interx

```

### Test AICc

```{r}
models <- list(mo_model_add, mo_model_interx) 
model.names <- c("mo_model_add", "mo_model_interx")

aictab(cand.set = models, modnames = model.names)

```

> Based on AICc test, the additive model is a better fit.


### Check assumptions {.tabset}
#### Interactive model

```{r, echo = FALSE}
# 1. Residuals-Fit plot check that variance is constant across groups
# --> Want: straight line; Avoid: funnel-like shape 
# 2. QQ plot checks for normality in the residuals
# --> Want: points fitting closely to the diagonal line
# https://mgimond.github.io/Stats-in-R/ANOVA.html

# Normality
OP <- par(mfrow=c(1,2))
plot(mo_model_interx, 1:2)
par(OP)

# Homogeneity of Variance
leveneTest(mumax~Temp*Nutrient, data = dat_mo)
```

> Data meets assumptions for the most part.


#### Additive model

```{r, echo = FALSE}
# plot(mo_model_interx, 1)
# plot(mo_model_interx, 2)
OP <- par(mfrow=c(1,2))
plot(mo_model_add, 1:2)
par(OP)
```




### Run ANOVA {.tabset}
#### Interactive ANOVA
```{r}
anova(mo_model_interx)
```

> Temp and nutrients do not have a significant effect on mumax; however, there is a significant interaction effect between them (p < 0.1). 


#### Additive ANOVA
```{r}
anova(mo_model_add)
```

> There is no significant effect of temp and nutrients on mumax.


# Average growth rate 
in progress...


#endscript