---
title: "Desmid Mini Project - ANOVA"
author: "Kristine Gesulga"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    code_folding: "hide" 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, include = FALSE, warnings = FALSE}

rm(list=ls())

# Load libraries
library(tidyverse) # lots of useful functions!
library(AICcmodavg) # to test AICc
library(car) # for Levene's Test

# Load data
raw_dat <- read_csv("data/1b_Growth_community_KG_edit2_mumax.csv")
glimpse(raw_dat)

```
# ANOVA
## Maximum growth rate
Question 1a: Does temperature and/or nutrients have a significant effect on maximum growth rate (mumax)?<br>
Question 1b: Do temperature and nutrients have an interactive effect?

### Meso community

<br>

#### Tidy data
<details>
  <summary>Click for details</summary>
```{r, message = FALSE}
dat_meso <- raw_dat %>% 
  filter(community == "meso") %>% 
  select(mumax, Temp, Nutrient)
glimpse(dat_meso) # response variable = mumax; explanatory variable = temperature / nutrient

# Convert Temperature + Nutrient treatments to factors (ANOVA test likes it in this format)
dat_meso$Temp <- as.factor(dat_meso$Temp)
dat_meso$Nutrient <- as.factor(dat_meso$Nutrient)
# Sanity check
table(dat_meso$Temp, dat_meso$Nutrient)
```
</details> 

<br>

#### Visualise the data

```{r}
# Plot
ggplot(dat_meso, aes(x = Nutrient, y = mumax, color = Temp)) +
  geom_boxplot()+
  labs(title = "Meso community")
```

<br>

#### Fit model
<details>
  <summary>Click for details</summary>
```{r, message = FALSE}
# Additive model
meso_model_add <- lm(mumax ~ Temp + Nutrient, data = dat_meso)
meso_model_add

# Interactive model
meso_model_interx <- lm(mumax ~ Temp*Nutrient, data = dat_meso)
meso_model_interx

```
</details>

<br>

#### Test AICc

```{r}
models <- list(meso_model_add, meso_model_interx) 
model.names <- c("meso_model_add", "meso_model_interx")

aictab(cand.set = models, modnames = model.names)

```


Based on AICc test, the additive model is a better fit.

<br>

#### Check assumptions {.tabset}
##### Interactive model

```{r}
# 1. Residuals-Fit plot check that variance is constant across groups
# --> Want: straight line; Avoid: funnel-like shape 
# 2. QQ plot checks for normality in the residuals
# --> Want: points fitting closely to the diagonal line
# https://mgimond.github.io/Stats-in-R/ANOVA.html

# Normality
OP <- par(mfrow=c(1,2))
plot(meso_model_interx, 1:2)
par(OP)

# Homogeneity of Variance
leveneTest(mumax~Temp*Nutrient, data = dat_meso)
```

Data meets assumptions for the most part.

<br>

##### Additive model
```{r}
# plot(meso_model_interx, 1)
# plot(meso_model_interx, 2)
OP <- par(mfrow=c(1,2))
plot(meso_model_add, 1:2)
par(OP)
```

<br>

#### Run ANOVA {.tabset}
##### Interactive ANOVA
```{r}
anova(meso_model_interx)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 2px; padding: 5px;}
</style>
<div class = "blue">

There is no significant effect of temp and nutrients on mumax or interactive effects.

</div>

##### Additive ANOVA
```{r}
anova(meso_model_add)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 2px; padding: 5px;}
</style>
<div class = "blue">

There is no significant effect of temp and nutrients on mumax.

</div>


<br>

### Meso-oligo community

<br>

#### Tidy data

<details>
  <summary>Click for details</summary>
```{r}
dat_mo <- raw_dat %>% 
  filter(community == "meso-oligo") %>% 
  select(mumax, Temp, Nutrient)
glimpse(dat_mo) # response variable = mumax; explanatory variable = temperature / nutrient

#### Convert Temperature + Nutrient treatments to factors (ANOVA test likes it in this format)
dat_mo$Temp <- as.factor(dat_mo$Temp)
dat_mo$Nutrient <- as.factor(dat_mo$Nutrient)
#### Sanity check
table(dat_mo$Temp, dat_mo$Nutrient)
```
</details> 

<br>

#### Visualise the data

```{r}
#### Plot
ggplot(dat_mo, aes(x = Nutrient, y = mumax, color = Temp)) +
  geom_boxplot()+
  labs(title = "Meso-oligo community")
```

<br>

#### Fit model
<details>
  <summary>Click for details</summary>
```{r}
# Additive model
mo_model_add <- lm(mumax ~ Temp + Nutrient, data = dat_mo)
mo_model_add

# Interactive model
mo_model_interx <- lm(mumax ~ Temp*Nutrient, data = dat_mo)
mo_model_interx

```
</details>

<br>

#### Test AICc

```{r}
models <- list(mo_model_add, mo_model_interx) 
model.names <- c("mo_model_add", "mo_model_interx")

aictab(cand.set = models, modnames = model.names)

```

Based on AICc test, the additive model is a better fit.

<br>

#### Check assumptions {.tabset}
##### Interactive model

```{r}
# Normality
OP <- par(mfrow=c(1,2))
plot(mo_model_interx, 1:2)
par(OP)

# Homogeneity of Variance
leveneTest(mumax~Temp*Nutrient, data = dat_mo)
```

Data meets assumptions for the most part.

<br>

##### Additive model

```{r}
OP <- par(mfrow=c(1,2))
plot(mo_model_add, 1:2)
par(OP)
```

<br>

#### Run ANOVA {.tabset}
##### Interactive ANOVA
```{r}
anova(mo_model_interx)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 2px; padding: 5px;}
</style>
<div class = "blue">

Temp and nutrients do not have a significant effect on mumax; however, there is a significant interaction effect between them <br>(p < 0.1). 

</div>
##### Additive ANOVA
```{r}
anova(mo_model_add)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 2px; padding: 5px;}
</style>
<div class = "blue">

There is no significant effect of temp and nutrients on mumax.

</div>

<br>

## Average growth rate 
in progress...

***

# Session Info
<details>
  <summary>Click for details</summary>
```{r}
sessionInfo()
```
</summary>
#endscript